\textbf{SPECFEM3D}\\ \textbf{Cartesian}\\ \textbf{User Manual}

### 2022-11-30 doc/USER_MANUAL/03_mesh_generation.tex

In each example subdirectory you will find a \texttt{README} file,
which explains in a step-by-step tutorial the workflow for the example.
Please feel free to contribute your own example to this package through the github repository!

@@

{\it IMPORTANT:} In order to correctly set up GEOCUBIT and run the examples, please read the file called \texttt{EXAMPLES/README};
in particular, please make sure you correctly set up the Python paths as indicated in that file.

@@

Once the geometric model volumes in CUBIT are meshed, you prepare
the model for exportation with the definition of material blocks and
boundary surfaces. Thus, prior to exporting the mesh, you need to
define blocks specifying the materials and absorbing boundaries in
CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
if the mesh meets some conditions or manually, following the block
convention:
\begin{description}
\item [{material\_name}] Each material should have a specific block defined
by a unique name. The name convention of the material is to start
with either \textbf{'elastic'} or \textbf{'acoustic'}. It must be then followed by a
unique identifier, e.g. \textbf{'elastic 1'}, \textbf{'elastic 2'}, etc. The additional
attributes to the block define the material description.

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. You can either specify a single Q value,
in which case it will be assumed to be pure shear attenuation $Q_{\mu}$, or two
separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other see doc/note\_on\_Qkappa\_versus\_Qp.pdf and utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90.

@@

Note that this material block has to be defined using all the volumes
which belong to this elastic material. For volumes belonging to another,
different material, you will need to define a new material block.

@@

Optionally, instead of specifying for each surface at the boundaries
a single block like mentioned above, you can also specify a single
block for all boundary surfaces and name it as one of the absorbing
blocks above, e.g. 'face\_abs\_xmin'.

@@

Note that it does not matter whether an element belongs to a Xmin
or to a Xmax CPML, the flag is the same in both cases; the same is
true for Ymin and Ymax, and also Zmin and Zmax.

When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
format, i.e., if you have existing \texttt{nodes\_coords\_file}
and \texttt{mesh\_file} files but do not know
how to assign CPML flags to them, we have created a small serial Fortran
program that will do that automatically for you, i.e., which will
create the \texttt{absorbing\_cpml\_file} for
you. That program is: \newline
\texttt{utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90},\newline
and a small Makefile is provided in that directory (utils/CPML).

IMPORTANT: it is your responsibility to make sure that in the input
CUBIT (or similar) mesh that this code will read in SPECFEM3D format
from files \texttt{nodes\_coords\_file} and \texttt{mesh\_file}
you have created layers of elements that constitute a layer of constant
thickness aligned with the coordinate grid axes (X, Y and/or Z), so
that this code can assign CPML flags to them. This code does NOT check
that (because it cannot, in any easy way). The mesh inside these CPML
layers does not need to be structured nor regular, any non-structured
mesh is fine as long as it has flat PML inner and outer faces, parallel
to the axes, and thus of a constant thickness. The thickness can be
different for the X, Y and Z sides. But for X it must not vary, for
Y it must not vary, and for Z it must not vary. If you do not know
the exact thickness, you can use a slightly LARGER value in this code
(say 2\% to 5\% more) and this code will fix that and will adjust
it; never use a SMALLER value otherwise this code will miss some CPML
elements.

Note: in a future release we will remove the constraint of having
CPML layers aligned with the coordinate axes; we will allow for meshes
that are titled by any constant angle in the horizontal plane. However
this is not implemented yet.

Note: in the case of fluid-solid layers in contact, the fluid-solid
interface currently needs to be flat and horizontal inside the CPML
layer (i.e., bathymetry should become flat and horizontal when entering
the CPML); this (small) constraint will probably remain in the code
for a while because it makes fluid-solid matching inside the CPML
much easier.

@@

where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
an identifier for anisotropic models.
Note that both \texttt{Qkappa} and \texttt{Qmu} are ignored
by the code unless \texttt{ATTENUATION} is set. If you want a model
with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
\texttt{.false.} in the \texttt{DATA/Par\_file} and set \texttt{Qmu} to
9999 here. If you want a model with no \texttt{Qkappa} attenuation, set \texttt{Qkappa} to 9999 here.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.

@@

\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
and Section \ref{sec:Using-tomographic} `Using external tomographic
Earth models' for further details.
\end{itemize}
\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
coordinates of the mesh element corners.
\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
elements can have 8 or 27 nodes.\newline
 See picture \texttt{doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg}
to see\newline
 in which (standard) order the points must be cited. In the case of
8 nodes, just include the first 8 points.
\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
connectivity or \newline
 the surface connectivity of the absorbing boundary surface at the
top (Zmax), \newline
 depending on whether the top surface is defined as free or absorbing
(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \newline
 You should put both the surface of acoustic regions and of elastic
regions in that file; that is, list all the element faces that constitute
the surface of the model in that file.
\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
of the absorbing boundary surface at Xmax \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
of the absorbing boundary surface at Xmin \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
of the absorbing boundary surface at Ymax \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
of the absorbing boundary surface at Ymin \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
of the absorbing boundary surface at the bottom (Zmin) \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\end{description}
These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
to load-balance the mesh. Please see the next section for further
details.

@@

The quality of the mesh may be inspected more precisely based upon
the serial code in the file \texttt{check\_mesh\_quality\_}~\newline
 \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
Running this code is optional because no information needed by the
solver is generated.

Prior to running and compiling this code, you have to export your
mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
block IDs belonging to volumes in order to check the quality of the
hexahedral elements. You also have to determine a number of parameters
of your mesh, such as the number of nodes and number of elements and
modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.

@@

The SPECFEM3D Cartesian software package performs large scale simulations
in a parallel 'Single Process Multiple Data' way. The spectral-element
mesh created with CUBIT needs to be distributed on the processors.
This partitioning is executed once and for all prior to the execution
of the solver so it is referred to as a static mapping.

@@

\noindent
When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
  make xdecompose_mesh
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

@@

\noindent
To run the
internal mesher for a regional simulation, the following parameters need to be
set in the \texttt{\small Mesh\_Par\_file}:
%
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when
\texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.

@@

From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\newline
The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
which means that current UTM northing at a given point can be 200+ meters different from the old one.
For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms, \cite{KoLiTrSuStSh04} determined
that a $\nexxi=288$ run for a 516 km x 507 km mesh
is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation with a mesh extent $\Delta$ (in km)
will be accurate approximately to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=2. \times \frac{288}{\nexxi} \times \frac{\Delta}{500km}\label{eq:shortest_period}
\end{equation}
This estimate also depends on the velocity model. Therefore, the binary \texttt{xgenerate\_databases}
calculates a minimum period resolved by the chosen mesh discretization and velocity model,
output into the file \texttt{OUTPUT\_FILES/output\_generate\_databases.txt}.

 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest this value as it represents a fair trade-off between element deformations to adapt to
topographic variations and numerical accuracy.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).

>>>>

In each example subdirectory you will find a \texttt{README} file,
which explains in a step-by-step tutorial the workflow for the example.
Please feel free to contribute your own example to this package through the github repository!\newline

@@

{\it IMPORTANT:} In order to correctly set up GEOCUBIT and run the examples, please read the file called \texttt{EXAMPLES/README};
in particular, please make sure you correctly set up the Python paths as indicated in that file.\newline

@@

Once the geometric model volumes in CUBIT are meshed, you prepare
the model for exportation with the definition of material blocks and
boundary surfaces. Thus, prior to exporting the mesh, you need to
define blocks specifying the materials and absorbing boundaries in
CUBIT. This process could be done automatically using the script \texttt{run\_cubit2specfem3d.py}
if the mesh meets some conditions or manually, following the block
convention:
\begin{description}
\item [{material\_name}] Each material should have a specific block defined
by a unique name. The name convention of the material is to start
with either \textbf{'elastic'} or \textbf{'acoustic'}. It must be then followed by a
unique identifier, e.g. \textbf{'elastic 1'}, \textbf{'elastic 2'}, etc. The additional
attributes to the block define the material description.\newline

For an elastic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{Vs}] S-wave speed of the material (given in m/s).
\item [{rho}] density of the material (given in kg/m$^{3}$).
\item [{Q}] quality factor to use in case of a simulation with attenuation
turned on. It should be between 1 and 9000. In case no attenuation
information is available, it can be set to zero. You can either specify a single Q value,
in which case it will be assumed to be pure shear attenuation $Q_{\mu}$, or two
separate values for bulk and shear attenuation, $Q_{\kappa}$ and $Q_{\mu}$ respectively.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp. To convert one to the other
see \texttt{doc/note\_on\_Qkappa\_versus\_Qp.pdf} and\newline
\texttt{utils/attenuation/conversion\_from\_Qkappa\_Qmu\_to\_Qp\_Qs\_from\_Dahlen\_Tromp\_959\_960.f90}.\newline

@@

Note that this material block has to be defined using all the volumes
which belong to this elastic material. For volumes belonging to another,
different material, you will need to define a new material block.\newline

@@

Optionally, instead of specifying for each surface at the boundaries
a single block like mentioned above, you can also specify a single
block for all boundary surfaces and name it as one of the absorbing
blocks above, e.g. 'face\_abs\_xmin'.\newline

@@

Note that it does not matter whether an element belongs to a Xmin
or to a Xmax CPML, the flag is the same in both cases; the same is
true for Ymin and Ymax, and also Zmin and Zmax.\newline

When you have an existing CUBIT (or similar) mesh stored in SPECFEM3D
format, i.e., if you have existing \texttt{nodes\_coords\_file}
and \texttt{mesh\_file} files but do not know
how to assign CPML flags to them, we have created a small serial Fortran
program that will do that automatically for you, i.e., which will
create the \texttt{absorbing\_cpml\_file} for
you. That program is: \newline
\texttt{utils/CPML/convert\_external\_layers\_of\_a\_given\_mesh\_to\_CPML\_layers.f90},\newline
and a small Makefile is provided in that directory (utils/CPML).\newline

IMPORTANT: it is your responsibility to make sure that in the input
CUBIT (or similar) mesh that this code will read in SPECFEM3D format
from files \texttt{nodes\_coords\_file} and \texttt{mesh\_file}
you have created layers of elements that constitute a layer of constant
thickness aligned with the coordinate grid axes (X, Y and/or Z), so
that this code can assign CPML flags to them. This code does NOT check
that (because it cannot, in any easy way). The mesh inside these CPML
layers does not need to be structured nor regular, any non-structured
mesh is fine as long as it has flat PML inner and outer faces, parallel
to the axes, and thus of a constant thickness. The thickness can be
different for the X, Y and Z sides. But for X it must not vary, for
Y it must not vary, and for Z it must not vary. If you do not know
the exact thickness, you can use a slightly LARGER value in this code
(say 2\% to 5\% more) and this code will fix that and will adjust
it; never use a SMALLER value otherwise this code will miss some CPML
elements.\newline

Note: in a future release we will remove the constraint of having
CPML layers aligned with the coordinate axes; we will allow for meshes
that are titled by any constant angle in the horizontal plane. However
this is not implemented yet.\newline

Note: in the case of fluid-solid layers in contact, the fluid-solid
interface currently needs to be flat and horizontal inside the CPML
layer (i.e., bathymetry should become flat and horizontal when entering
the CPML); this (small) constraint will probably remain in the code
for a while because it makes fluid-solid matching inside the CPML
much easier.\newline

@@

where \texttt{\textbf{domain\_ID}}\textbf{ is 1 for acoustic and 2
for elastic or viscoelastic materials,} \texttt{material\_ID} a unique
identifier, \texttt{rho} the density in $kg\, m^{-3}$, \texttt{vp}
the P-wave speed in $m\, s^{-1}$, \texttt{vs} the S-wave speed in
$m\, s^{-1}$, \texttt{Q} the quality factor and \texttt{anisotropy\_flag}
an identifier for anisotropic models.
Note that both \texttt{Qkappa} and \texttt{Qmu} are ignored
by the code unless \texttt{ATTENUATION} is set. If you want a model
with no \texttt{Qmu} attenuation, both set \texttt{ATTENUATION} to
\texttt{.false.} in the \texttt{DATA/Par\_file} and set \texttt{Qmu} to
9999 here. If you want a model with no \texttt{Qkappa} attenuation, set \texttt{Qkappa} to 9999 here.
Note that Qmu is always equal to Qs, but Qkappa is in general not equal to Qp.\newline

@@

\item For tomographic velocity models, please read Chapter \ref{cha:-Changing-the}
and Section \ref{sec:Using-tomographic} `Using external tomographic
Earth models' for further details.
\end{itemize}
\item [{nodes\_coords\_file}] Contains the point locations in Cartesian
coordinates of the mesh element corners.
\item [{mesh\_file}] Contains the mesh element connectivity. The hexahedral
elements can have 8 or 27 nodes.\newline
 See picture \texttt{doc/mesh\_numbering\_convention/numbering\_convention\_27\_nodes.jpg}
to see\newline
 in which (standard) order the points must be cited. In the case of
8 nodes, just include the first 8 points.
\item [{free\_or\_absorbing\_surface\_file\_zmax}] Contains the free surface
connectivity or \newline
 the surface connectivity of the absorbing boundary surface at the
top (Zmax), \newline
 depending on whether the top surface is defined as free or absorbing
(\texttt{STACEY\_INSTEAD\_OF\_FREE\_SURFACE} in \texttt{DATA/Par\_file}). \newline
 You should put both the surface of acoustic regions and of elastic
regions in that file; that is, list all the element faces that constitute
the surface of the model in that file.
\item [{absorbing\_surface\_file\_xmax}] Contains the surface connectivity
of the absorbing boundary surface at Xmax \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_xmin}] Contains the surface connectivity
of the absorbing boundary surface at Xmin \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_ymax}] Contains the surface connectivity
of the absorbing boundary surface at Ymax \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_ymin}] Contains the surface connectivity
of the absorbing boundary surface at Ymin \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\item [{absorbing\_surface\_file\_bottom}] Contains the surface connectivity
of the absorbing boundary surface at the bottom (Zmin) \newline
(also needed in the case of C-PML absorbing conditions, in order for
the code to be able to impose Dirichlet conditions on their outer
edge).
\end{description}
These mesh files are needed as input files for the partitioner \texttt{xdecompose\_mesh}
to load-balance the mesh. Please see the next section for further
details.\newline

@@

The quality of the mesh may be inspected more precisely based upon
the serial code in the file \texttt{check\_mesh\_quality\_}~\newline
 \texttt{CUBIT\_Abaqus.f90} located in the directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.
Running this code is optional because no information needed by the
solver is generated.\newline

Prior to running and compiling this code, you have to export your
mesh in CUBIT to an ABAQUS (.inp) format. For example, export mesh
block IDs belonging to volumes in order to check the quality of the
hexahedral elements. You also have to determine a number of parameters
of your mesh, such as the number of nodes and number of elements and
modify the header of the \texttt{check\_mesh\_quality\_CUBIT\_Abaqus.f90}
source file in directory \texttt{src/check\_mesh\_quality\_CUBIT\_Abaqus/}.\newline

@@

The SPECFEM3D Cartesian software package performs large scale simulations
in a parallel 'Single Process Multiple Data' way. The spectral-element
mesh created with CUBIT needs to be distributed on the processors.
This partitioning is executed once and for all prior to the execution
of the solver so it is referred to as a static mapping.\newline

@@

\noindent
When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
  make xdecompose_mesh
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.\newline

@@

\noindent
To run the
internal mesher for a regional simulation, the following parameters need to be
set in the \texttt{\small Mesh\_Par\_file}:
%
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when
\texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.\newline

@@

From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\newline
The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
which means that current UTM northing at a given point can be 200+ meters different from the old one.
For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.\newline

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms, \cite{KoLiTrSuStSh04} determined
that a $\nexxi=288$ run for a 516 km x 507 km mesh
is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation with a mesh extent $\Delta$ (in km)
will be accurate approximately to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=2. \times \frac{288}{\nexxi} \times \frac{\Delta}{500km}\label{eq:shortest_period}
\end{equation}
This estimate also depends on the velocity model. Therefore, the binary \texttt{xgenerate\_databases}
calculates a minimum period resolved by the chosen mesh discretization and velocity model,
output into the file \texttt{OUTPUT\_FILES/output\_generate\_databases.txt}.\newline

 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{setup/constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest this value as it represents a fair trade-off between element deformations to adapt to
topographic variations and numerical accuracy.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set at least to \texttt{1} if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}.
Multiple mesh doublings can be chosen, for which each an \texttt{NZ\_DOUBLING\_**} entry must be given.
By default, we only provide two possible entries in the \texttt{Mesh\_Par\_file}.
For higher numbers of doubling layers, additional entries must be added.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.false.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}). Doubling layers must be at least \texttt{2} layers apart.
The layer count starts from the bottom layer.
More entries must be listed by the user if \texttt{NDOUBLINGS} is larger than \texttt{2}.
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).\newline
